\begin{abstract}
  Systems designed for mobile and Internet-of-things (IoT) must
  deliver performance to interactive applications (that monitor the
  environment and respond to users) and they must conserve resources
  to extend battery life.  Meeting these conflicting goals is
  complicated by two issues: (1) hardware heterogeneity leading to
  complicated optimization spaces and (2) dynamic operating conditions
  including unpredictable changes in available resources and
  application behavior.  Machine learning techniques handle
  complicated optimization spaces, but do not incorporate models of
  system dynamics; optimal control techniques provide formal
  guarantees of dynamic behavior, but cannot handle non-linear system
  models.  In this paper, we propose a combination of learning and
  control techniques to meet mobile system's performance requirements
  in unpredictable environments.  Specifically, we combine a
  hierarchical Bayesian model (HBM) with a lightweight control system
  (LCS).  The HBM runs on a remote server, aggregating data from many
  separate mobile systems to learn customized models of
  performance/power tradeoffs.  The LCS runs locally on the mobile
  system and tunes resource usage to meet performance requirements
  efficiently.  We implement both the HBM and LCS and test their
  ability to learn and control ARM big.LITTLE systems.  Compared to
  existing learning and control methods, our proposed combined system
  delivers more reliable performance -- only XX\% error compared to
  YY\% for learning and ZZ\% for control -- and higher energy
  efficiency -- achieving ZZ\% of optimal and improving by XX\% over
  learning and YY\% over control.  When used in a multi-application
  scenario these numbers improve: xx\% error compared to yy\% for
  learning and zz\% for control and improvements of xx\% and zz\%,
  respectively, in energy efficiency.
\end{abstract}
